---
title: "Example Timings"
author: "Robert M Flight"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Timings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose

The goal of `importedPackageTimings` is to help developers determine if any of the
R packages their package depends on make loading their package slow.

To accompmlish this, it uses independent R sessions from the `future` package
to time how long it takes to load each of the packages listed in the `Imports`
and `Depends` fields of the package in question. Although it will take a long
time because it only uses a single core at a time (the only way I could get
reliable timings), the times seem to be reliable.

## Installation

Currently, `dependencyTimings` only exists on Github, so install it with:

```{r install, eval=FALSE}
remotes::install_github("rmflight/dependencyTimings")
```

## Example

For example, lets look at a Bioconductor package I've seen take a long time to
load, `xcms`.

```{r example, eval=FALSE}
# not run
library(furrr)
plan(multiprocess)
library(importedPackageTimings)
xcms_time = imported_timings("xcms")
```

The package provides two types of timings, the time required for the dependency
to load (type = `pkg`), and then the time required for the package to load after
the dependency (type = `after`).

```{r show_table, results = 'asis'}
library(importedPackageTimings)
data("xcms_time")
knitr::kable(head(dplyr::select(xcms_time, -timings)))
```

We can use the `pkg` entries to see which dependencies actually take a long time
to load, possibly contributing to the long load time of our package in question.

```{r show_pkg}
library(ggplot2)
ggplot(dplyr::filter(xcms_time, type %in% "pkg"), 
       aes(x = min / 1e9, y = package)) + 
  geom_point()
```

From this plot, we can see that `MSnbase` looks like it is taking the longest to
load outside of `xcms` itself. 

We can use the `after` entries to see which dependencies after loading have the
smallest time to load our package in question, which implies they may be the
culprit causing long load times.

```{r show_after}
ggplot(dplyr::filter(xcms_time, type %in% "after", which %in% "import"),
       aes(x = min / 1e9, y = package)) +
  geom_point()
```
